---
title: "Calculating Euclidean and along-river distance between sites"
author: "Sarah J. Hart"
email: "sarah.hart@colostate.edu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
   officedown::rdocx_document:
     reference_docx: WordTemplate.docx
bibliography: references.bib
csl: "`r here::here('ecology.csl')`"
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  results = FALSE,
	message = FALSE,
	warning = FALSE,
	progress = FALSE,
	cache = FALSE,
	dpi = 300
)

set.seed(513)
options(repos = c(CRAN = "http://cran.rstudio.com"))
options(timeout=60*30) #timeout downloads that last longer than 30 minutes

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  devtools,
  knitr, # markdown documents
  flextable, # plot tables
  bookdown, # figure numbering in markdown
  here, # easy file structure
  tidyverse, # data manipulation
  sf, # spatial data (new pkg)
  archive, #unzipping .7z files
  readxl, # read excel files
  units, # support for measurement units
  sfnetworks #spatialneworks
)

# Set custom plotting theme
theme_new <- function(base_size = 9,base_family = "Helvetica"){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.line.x = element_line(color="black", linewidth = 0.25),
      axis.line.y = element_line(color="black", linewidth = 0.25),
      axis.title = element_text(size = 9),
      axis.text = element_text(colour="black", size=8),
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      plot.background = element_rect(fill = NA, colour = NA),
      panel.border = element_rect(fill = NA, colour = NA),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 9)
      
    )
}
theme_set(theme_new())

set_flextable_defaults(
  font.family="Times", 
  font.size=12,
  line_spacing=1,
  padding.bottom=1,
  padding.top=1,
  text.align='center')

# Set directory structure for project
dir.create(here("Data"), showWarnings = FALSE)
dir.create(here("Data", "Spatial"), showWarnings = FALSE)
dir.create(here("Results"), showWarnings = FALSE)
dir.create(here("Results", "Figures"), showWarnings = FALSE)
```

# Data

```{r}
sites <- read_csv(here("Data", "Jordan_sites.csv")) 
sites.sf <- st_as_sf(sites, coords = c("UTM_easting", "UTM_northing"),  crs = 32613) 
st_write(sites.sf, here("Data", "Spatial", "sites.shp"), append=F)
```

## Download data from NHDPlusV2

We downloaded high resultion hydrography data from the USGS NHDPlus Version 2 using the *nhdplusTools* package [@nhdplusTools].

```{r}
library(nhdplusTools)
dir.create(here("Data", "Spatial", "NHDPlusV2"))
studyarea <- sites.sf %>% st_buffer(100) %>% st_union() # create buffer around sites and union to create study area
huc06 <- get_huc(AOI = studyarea, id = NULL, t_srs = NULL, buffer = 0.5, type = "huc06") # this downloads all of teh data for the Rio Grande Watershed
huc12 <- get_huc(AOI = studyarea, id = NULL, t_srs = NULL, buffer = 0.5, type = "huc12")

download_nhdplushr(nhd_dir=here("Data", "Spatial", "NHDPlusV2"), hu_list=huc06$huc6, download_files = TRUE, archive = FALSE)

nhdplus.gdbs <- list.files(here("Data", "Spatial", "NHDPlusV2", "13"), pattern=".gdb", full.names=T)
nhdplus.flow <- lapply(nhdplus.gdbs, function(x){read_sf(x, layer="NHDFlowline")}) %>% dplyr::bind_rows()
nhdplus.wbd8 <-  lapply(nhdplus.gdbs, function(x){read_sf(x, layer="WBDHU8")}) %>% dplyr::bind_rows() 
st_write(nhdplus.wbd8, here("Data", "Spatial", "NHDPlusV2", "NHDwbd8.shp"), append=F)
st_write(nhdplus.flow, here("Data", "Spatial", "NHDPlusV2", "NHDFlowline.shp"), append=F)
```

We the manually digitized one stream that was not included in the NHDPlusV2 dataset using QGIS.

```{r import_Data}
nhdplus.wbd8.sub <- nhdplus.wbd8 %>% filter(Name=="Alamosa-Trinchera")
streams <- st_read(here("Data", "Spatial", "NHDPlusV2", "NHDFlowline-edit.shp")) %>% st_crop(nhdplus.wbd8.sub) # select only streams in Alamosa-Trinchera watershed
```

# Calculate distance between point

## Euclidean distance

We calculated the Euclidean distance between sites using the *sf* package [@sf].

```{r Calc_Euclidean}
sites.dist.euclidean <- st_distance(sites.sf) %>% as.data.frame() %>% drop_units() # calculate euclidean distance between sites in meters and drop units so that data can be saved as csv
colnames(sites.dist.euclidean) <- sites$Site # rename columns to site names
row.names(sites.dist.euclidean) <- sites$Site # rename rows to site names
write.csv(sites.dist.euclidean, here("Results", "sites-euclidean.csv")) # write data to csv file 
st_write(sites.sf, here("Data", "Spatial", "sites.shp"), append=F) # write data to shapefile for visualization
```

## Distance along rivers

We calculated the distance between sites along rivers using the *sfnetworks* package [@sfnetworks].

```{r Calc_Stream}
streams.sub <- streams %>% 
  st_zm() %>% # drop M dimension
  st_transform(st_crs(sites.sf)) # reproject to UTM 13N
 
streams.sub <- streams.sub  %>% 
  st_union() %>% # dissolve lines and create nodes for every end point and intersection point
  st_cast("LINESTRING") %>% # convert to linestring
  as_sfnetwork(directed = FALSE) %>%  # create network object 
  st_network_blend(sites.sf, tolerance=100)  %>% # snap sites to network
  activate(edges) %>% # activate edges
  st_as_sf() %>% # convert to sf object
  st_write(here("Data", "Spatial", "edges.shp"), append=F) # write edges to file for visualization

river.dist <- st_read(here("Data", "Spatial", "edges.shp")) %>% 
  st_union() %>% # dissolve lines and create nodes for every end point and intersection point
  st_cast("LINESTRING") %>% # convert to linestring
  as_sfnetwork(directed = FALSE) %>%  # create network object
  st_network_cost(from=sites.sf, to=sites.sf, direction="all")  # calculate distance along rivers between sites

river.dist <- data.frame(river.dist, row.names=sites.sf$Site) # convert to data frame
colnames(river.dist) <- sites.sf$Site # rename columns to site names
write.csv(river.dist, here("Results", "sites-river.csv")) # write data to csv file
```

# References
